### 2023-04-02 hw1/paper/main.tex
update paper

%\documentclass[conference]{IEEEtran}
%\IEEEoverridecommandlockouts
%\documentclass[sigconf]{acmart}
%\let\Bbbk\relax %% fix bug
\documentclass[twocolumn,10pt]{article}
\usepackage[utf8]{inputenc}

% =======================
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amstext}
\usepackage{amsopn}
%\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}

\usepackage{textcomp}

\usepackage{boxedminipage}
\usepackage{enumerate}
\usepackage{multirow}
\usepackage{url}
\usepackage{times}
\usepackage{version}
% \usepackage[pdftex]{graphicx}
\usepackage{epsfig}
\usepackage{epsf}
%\usepackage{graphics}
\usepackage{caption}
\usepackage{subfigure}
\usepackage{algorithm}
\usepackage{algpseudocode}
%\PassOptionsToPackage{bookmarks={false}}{hyperref}
%%%%%%%%%%%%
\usepackage{comment}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage{dblfloatfix}
% ==========================
%\usepackage[a4paper, margin = 2.5cm]{geometry}

\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

\begin{document}

\title{The Accuracy of KNN, decision tree, random forest, SVM, neural network, naive Bayes classifier and PLA for early Prediction of Diabetes}

\author{Hsieh Cheng-Han}
\date{October 2022}
\maketitle

\section*{abstract}
  This work compares the accuracy of some classifiers for early the prediction of diabetes. More specifically, 
  the research compares the accuracy of k-nearest neighbors (KNN) algorithm, decision tree, random forest, 
  support vector machine (SVM), neural network, naive Bayes classifier, and PLA on the prediction of diabetes, which 
  the dataset is collected with eight features, times of pregnancy, concentration of glucose in blood, blood 
  pressure, skin thickness,  concentration of insulin in blood, body mass index (BMI), the value of diabetes 
  pedigree function and age.

  The result show that XXX is the most accurate on the prediction of diabetes.

\section{Introduction}
\label{sec:Introduction}
  Diabetes is a chronic disease which may cause many complications. There're lots of reason that can put a person 
  at the highly risk of having diabetes, such as age, obesity, lack of exercies, and more on. So many reasons 
  interweave together making the manual prediction on diabetes is nearly impossible. However, lots of works \cite{MUJUMDAR2019292} \cite{MAHBOOBALAM2019100204} \cite{10.3389/fgene.2018.00515}
  show that it is possible to have high accuracy by using machine learning techniques, such as random forest, 
  K-means clustering, neural network, and so on.

  By collecting the essential data of human body, prediction of diabetes can be turn into classification problem. 
  Imagine that an individual case with essential data is a point in hyperspace, if it is closer to the cluster
  having diabetes, this case is more likely to have diabetes in the future, otherwise, this case is more likely 
  healthy. But there are lots of machine learning techniques born to solve classification problem, it remains a 
  problem that which technique having the hightest accuracy on the prediction of diabetes.

  To find out which techniques is more suitable to predict diabetes, this work examines the diagnosis of diabetes 
  using KNN algorithm, decision tree, random forest, SVM, neural network naive Bayes classifier, and PLA.
\section{Related works}
\label{sec:Related works}
  \bf{k-nearest neighbors (KNN) classification algorithm}: The KNN classification algorithm is a supervised learning 
  method which is first developed by Fix and Hodges \cite{10.2307/1403797}. The idea of KNN is based on the idiom, 
  "birds of a feather flock together". By pick the $k$-nearest neighbors of a data point, the class label can be determined.

  The rough process of KNN is described below: suppose that there is a dataset which contain $N$ data point, denoted as 
  $(X_i,Y_i)$ where $X_i$ is the features of the i-th individuals data and $Y_i$ is the class label of it. Now 
  a data with unknown class label is given, denoted $(X, Y)$. By a preset distance function $d(P, Q)$, ordering the dataset 
  as $(X_{(1)}, Y_{(1)}), (X_{(2)}, Y_{(2)}), \cdots, (X_{(N)}, Y_{(N)})$ where $d(X_{(1)}, X)\leq d(X_{(2)}, X)\leq\cdots\leq d(X_{(N)}, X)$. 
  Pick the $k$-first class labels to determine the unknow class label, $Y$.

  \bf{decision tree}: Decision tree is a supervised learning method. Unlike KNN uses distance to determine the outcome, 
  decision tree uses a sequence of deicsion that maximize the information gain, which can distinguish the class label of 
  data as much as possible, to determine the outcome.  Given a dataset, $D$, which contain $N$ data and class label, the 
  construction of a decision tree can be described as below: suppose there are $M$ decisions, denoted $f_i$. A decision can 
  separate dataset $D$ into $m$ kinds, denote $D'_j$. The decision tree will adopt $\max_{f} G(D, f)=I(D)-\sum^m_{j=1}\frac{N'_j}{N}I(D'_j)$, 
  and then recursively construct the tree until the data in separated dataset have the same class label. 
  When a data with unknown class label comes, a decision tree determines recursively by the decision node until the leaf node. 
  The advantage of decision tree is fast, and the decision is clear. But the disadvantage is that it is very likely overfitting 
  and the structure of tree will become more complex with the more the class labels. To solve this problem, the following 
  techniques is developed:

  \bf{random forst}: Instead of a signle decision tree, random forest use lots of decision trees, which form a "forest". 
  The decision trees are constructed by random subset of dataset. The key differs random forest from decision tree is that 
  while decision trees consider all the possible feature splits, random forests only select a subset of those features, which 
  reduce the risk of overfitting, bias, and overall variance.

  \bf{support vector machine (SVM)}: 
  \bf{neural network}: 
  \bf{naive Bayes classifier}: 
  \bf{PLA}:

\section{KNN}

\section{SVM}

\section{neural network}

\section{Naive Bayes classifier}

\section{PLA}

\section{Experiment result}

\section{Conclusion}

\bibliographystyle{IEEEtran}
\bibliography{main}
\end{document}